Typing a linear \picalc

### 2020-03-26 paper/paper.tex
Minor fixes

@@

@@

\begin{figure}[h]
  \begin{mathpar}
    {\mprset{fraction={===}}
      \inferrule
      { }
      {\type{\N} : \Set}}

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    {n : \N}
    {\Var_n : \Set}}

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    {n : \N}
    {\Process_n : \Set}}
\end{mathpar}

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    { }
    {P \eqeq Q : \Set}}

@@

\begin{figure}[h]
\begin{mathpar}
  \mprset{flushleft}

  {\mprset{fraction={===}}
    \inferrule
    { }
    {\Rec : \Set}}

@@

  {\mprset{fraction={===}}
    \inferrule
    {P \, Q : \Process_n \\ r : \Rec}
    {P \eq{r} Q : \Set}}

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    {n : \N}
    {\Channel_n : \Set}}

@@

  {\mprset{fraction={===}}
    \inferrule
    {i : \Channel_n \\ P \; Q : \Process_n}
    {P \reduce{i} Q : \Set}}

@@

\begin{figure}[h]
\begin{equation}
\begin{split}
  \field{IDX}          &: \Set \\
  \field{\exists IDX}  &: \field{IDX} \\
  \field{CARRIER}      &: \field{IDX} \to \Set \\
  \field{QUANTIFIERS}  &: (i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
\end{split}
\end{equation}
\caption{Indexed set of partial commutative monoids}
\label{indexed-capabilities}
\end{figure}

@@

\todo{mention types are independent from capabilities}
\todo{add index to in-type}
\todo{update examples}

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    { }
    {\Type : \Set}
  }

@@

  \inferrule
  {t : \Type \\ \inferrule*{} {i : \Idx \\\\ x : \Carrier_i}}
  {\channel{t}{x} : \Type}
\end{mathpar}
\caption{This is...}
\end{figure}

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    {n : \N}
    {\PreCtx_n : \Set}
  }

  \inferrule
  {\gamma : \PreCtx_n \\ t : \Type}
  {\gamma , t : \PreCtx_{\suc n}}
\end{mathpar}
\caption{This is...}
\end{figure}

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    {n : \N}
    {\Idxs_n : \Set}}

  \inferrule
  {is : \Idxs_n \\ i : \Idx}
  {is , i : \Idxs_{\suc n}}

  {\mprset{fraction={===}}
    \inferrule
    {is : \Idxs_n}
    {\Ctx_{is} : \Set}}

  \inferrule
  {\Gamma : \Ctx_{is} \\ x : \Carrier_i}
  {\Gamma , x : \Ctx_{is , i}}
\end{mathpar}
\caption{This is...}
\end{figure}
two-layered approach: types on one hand, capabilities on the other
removing from context vs keeping in context but marking it used

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule{
      \gamma : \PreCtx_n \\
      \inferrule*{}{
        idxs : \Idxs_n \\\\
        \Gamma : \Ctx_{idxs}} \\
      i : Fin n
      t : \Type \\
      x : \Carrier_i \\
      \Delta : \Ctx_{is}}
    {\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta} : \Set}}

  \inferrule
  {\Gamma : \Ctx_{idxs} \\ y \; z : \Carrier_i \\ True (\constr{\cdot-compute} \; y \; z)}
  {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{zero}{t}{y}{\Gamma , z}}

  \inferrule
  {\Gamma : \Ctx_{idxs} \\ x : \Carrier_i \\ x' : \Carrier_j \\ \Delta : \Ctx_{idxs} \\\\
    loc_x : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
  {\constr{suc} \; loc_x : \contains{\gamma , t}{\Gamma , x'}{\suc i}{t}{x}{\Delta , x'}}
\end{mathpar}
\caption{This is...}
\end{figure}

@@

\begin{figure}[h]
\begin{mathpar}
  \mprset{flushleft}

  {\mprset{fraction={===}}
    \inferrule{
      \gamma : \PreCtx_n \\
      \inferrule*{}{
        idxs : \Idxs_n \\\\
        \Gamma : \Ctx_{idxs}} \\
      P : \Process_n \\
      \Delta : \Ctx_{idxs}}
    {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}}

@@

  \inferrule
  {t : \Type \\ x : \Carrier_i \\ y : \Carrier_j \\\\
   cont : \types{\gamma , \channel{t}{x}}{\Gamma , y}{P}{\Delta , \lz}}
  {\constr{chan} \; t \; x \; y \; cont : \types{\gamma}{\Gamma}{\new P}{\Delta}}

@@

  \inferrule
  {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
   r : \types{\gamma}{\Delta}{Q}{\Xi}}
  {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
\end{mathpar}
\caption{This is...}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject Reduction}\label{subject-reduction}

>>>>

% Less margins around figures
\usepackage{changepage}

@@

\newcommand{\stacked}[1]{\mprset{flushleft} \inferrule*{}{#1}}
\newcommand{\datatype}[2]{{\mprset{fraction={===}} \inferrule{#1}{#2}}}

@@

\begin{figure}[h]
  \begin{mathpar}
    \datatype
    { }
    {\type{\N} : \Set}

@@

\begin{figure}[h]
\begin{mathpar}
  \datatype
  {n : \N}
  {\Var_n : \Set}

@@

\begin{figure}[h]
\begin{mathpar}
  \datatype
  {n : \N}
  {\Process_n : \Set}
\end{mathpar}

@@

\begin{figure}[h]
\begin{mathpar}
  \datatype
  { }
  {P \eqeq Q : \Set}

@@

\begin{figure}[h]
\begin{mathpar}
  \datatype
  { }
  {\Rec : \Set}

@@

  \datatype
  {P \, Q : \Process_n \\ r : \Rec}
  {P \eq{r} Q : \Set}

@@

\begin{figure}[h]
\begin{mathpar}
  \datatype
  {n : \N}
  {\Channel_n : \Set}

@@

  \datatype
  {i : \Channel_n \\ P \; Q : \Process_n}
  {P \reduce{i} Q : \Set}

@@

\begin{figure}[h]
\begin{equation}
\begin{split}
  &\field{IDX}          &:{} &\Set \\
  &\field{\exists IDX}  &:{} &\field{IDX} \\
  &\field{CARRIER}      &:{} &\field{IDX} \to \Set \\
  &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
\end{split}
\end{equation}
\caption{Indexed set of partial commutative monoids}
\label{indexed-capabilities}
\end{figure}

@@

@@

In the linear \picalc{}, the capabilities of the variables in context need to be modified every time input or output occurs.
However, the types of the variables in context need not.
As neither capabilities nor their indices depend on types, we choose to keep these two kinds of information separate (Figure \ref{contexts}): a length-indexed context of types $\PreCtx_n$ on one side; a length-indexed context of indices $\Idxs_n$ on the other, with a context of capabilities $\Ctx_{idxs}$ on top.

\begin{figure}[h]
  \begin{mathpar}
    \datatype
        {n : \N}
        {\PreCtx_n : \Set}

        \inferrule
            {\gamma : \PreCtx_n \\ t : \Type}
            {\gamma , t : \PreCtx_{\suc n}}
  \end{mathpar}

  \begin{mathpar}
    \datatype
        {n : \N}
        {\Idxs_n : \Set}

        \inferrule
            {idxs : \Idxs_n \\ idx : \Idx}
            {idxs , idx : \Idxs_{\suc n}}

            \datatype
                {idxs : \Idxs_n}
                {\Ctx_{idxs} : \Set}

                \inferrule
                    {\Gamma : \Ctx_{idxs} \\ x : \Carrier_{idx}}
                    {\Gamma , x : \Ctx_{idxs , idx}}
  \end{mathpar}
  \caption{Length-indexed typing contexts; length-indexed index context; index-indexed capability contexts.}
  \label{contexts}
\end{figure}

We represent base types $\base{n}$ using a natural number $n$ that can be reified into a type by the host language.
Using natural numbers instead of type universes is convenient: there is no need to litter type types with universe polymorphism, and it provides an additional proof of inhabitance, needed for the $\new \PO_{\suc n} \eqeq \PO_n$ rule in subject congruence.

Channel types $\channel{t}{x}$ use $t$ as the type of data sent over the channel and $x$ as the capability of the data sent over the channel.
When data is received on a channel of type $\channel{t}{x}$, the type $t$ will be appended to the context of types $\PreCtx_n$ and the capability $x$ to the context of capabilities $\Ctx_{idxs}$.

\begin{figure}[h]
\begin{mathpar}
  \datatype
  { }
  {\Type : \Set}

@@

  \inferrule
  {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}}}
  {\channel{t}{x} : \Type}
\end{mathpar}
\caption{Types for a resource-aware \picalc{}.}
\label{types}
\end{figure}

@@

two-layered approach: types on one hand, capabilities on the other

\begin{figure}[h]
\begin{adjustwidth}{-1cm}{-1cm}
\centering
\begin{mathpar}
  \mprset{sep=0.5em}

  \datatype{
    \gamma : \PreCtx_n \\
    \stacked{
      idxs : \Idxs_n \\\\
      \Gamma : \Ctx_{idxs}} \\
    i : \Var_n \\
    t : \Type \\
    \stacked{
      idx : \Idx \\\\
      x : \Carrier_{idx}} \\
    \Delta : \Ctx_{idxs}}
  {\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta} : \Set}

  \inferrule
      {\gamma : \PreCtx_n \\
        \stacked{idxs : \Idxs_n \\\\ \Gamma : \Ctx_{idxs}}\\
        t : \Type \\
        \stacked{idx : \Idx \\\\ x \; y \; z : \Carrier_{idx}}\\
        \op{x}{y}{z}}
  {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{zero}{t}{y}{\Gamma , z}}

  \inferrule
      {\gamma : \PreCtx_n \\
        \stacked{idxs : \Idxs_n \\\\ \Gamma : \Ctx_{idxs}}\\
        t \; t' \; : Type \\
        \stacked{idx : \Idx \\\\ x : \Carrier_{idx}} \\
        \stacked{idx' : \Idx \\\\ x' : \Carrier_{idx'}} \\
        \Delta : \Ctx_{idxs} \\\\
    loc_x : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
  {\constr{suc} \; loc_x : \contains{\gamma , t'}{\Gamma , x'}{\suc i}{t}{x}{\Delta , x'}}
\end{mathpar}
\end{adjustwidth}
\caption{This is...}
\end{figure}

@@

\begin{figure}[h]
\begin{mathpar}
  \datatype{
    \gamma : \PreCtx_n \\
    \stacked{
      idxs : \Idxs_n \\\\
      \Gamma : \Ctx_{idxs}} \\
    P : \Process_n \\
    \Delta : \Ctx_{idxs}}
  {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}

@@

  \inferrule
  {t : \Type \\ x : \Carrier_{idx} \\ y : \Carrier_j \\\\
   cont : \types{\gamma , \channel{t}{x}}{\Gamma , y}{P}{\Delta , \lz}}
  {\constr{chan} \; t \; x \; y \; cont : \types{\gamma}{\Gamma}{\new P}{\Delta}}

@@

  \inferrule
  {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
   r : \types{\gamma}{\Delta}{Q}{\Xi}}
  {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
\end{mathpar}
\caption{This is...}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject Reduction}\label{subject-reduction}
